---
title: "Big Five Analysis"
output:
  github_document
---

This is an example showing how to generate various plots including radar charts, dendrograms, bar plots and PCA using dummy data of the Big Five Personality Test. The Big Five Personality Test is a free, open-source test consisting of 120 simple questions. You can take the test through this [link](https://bigfive-test.com/) and learn more about the test there. Here is the original publication by [Johnson, J. A. (2014)](https://doi.org/10.1016/j.jrp.2014.05.003).

## 1. Take a look at the `data`

The example data is available in `data.Rdata`. You can load the file with `load()`.

```{r, eval=FALSE}
load("data.Rdata")
```

It can also be loaded from the original google sheet with the `googlesheets4` package. The code is available in `1_load_googlesheet.R`.

```{r, eval=FALSE}
library(googlesheets4)
data <- read_sheet("https://docs.google.com/spreadsheets/d/1ZE0DmZfJcgQ_K_1jsta_OAVmADc8z5F_aLhJlEk7CZU/edit#gid=0")
```

After the dummy data is loaded into the R environment, we can take a look at it.

```{r echo = FALSE, results = 'axis'}
library(knitr)
load("data.Rdata")
knitr::kable(data[1:9,])
```

The dataset contains 9 samples. The 5 columns after `Sample no.` correspond to the 5 domains of the test: [`Neuroticism`](https://bigfive-test.com/neuroticism), [`Extraversion`](https://bigfive-test.com/extraversion), [`Openness To Experience`](https://bigfive-test.com/openness_to_experience), [`Agreeableness`](https://bigfive-test.com/agreeableness) and [`Conscientiousness`](https://bigfive-test.com/conscientiousness). 

Other columns are the 6 sub-categories under each domain. Detailed descriptions of the domains and the sub-categories can be found in the hyperlinks above.

## 2. Radar chart showing scores of each domain

You may wish to generate a radar chart with [`ggradar`](https://github.com/ricardo-bion/ggradar). In this exercise, the original code of `ggradar` has been modified. The modified codes are available in the folder `ggradar_edit`.

With the function `gen_radar()`, a radar plot with the 5 big items is generated. Below shows an example using sample no. 1. Aesthetics can be changed in the source code `1_gen_radar.R`. For simplicity, the 5 items are abbreviated with their initials. R script of this part is available in `2_gen_radar.R`.

```{r, eval=FALSE}
load("radar.Rdata")
radar_list_titleT[[1]]
```

```{r echo = FALSE, results = 'axis', fig.align='center'}
load("radar.Rdata")
radar_list_titleT[[1]]
```

Below shows the combined radar plots of all 9 samples. Each sample is assigned with a specific color.

```{r, eval=FALSE}
library(gridExtra)
grid.arrange(grobs = radar_list_titleT, nrow = 3, padding = unit(0, "line"))
```

```{r echo = FALSE, results = 'axis', fig.align='center'}
library(gridExtra)
grid.arrange(grobs = radar_list_titleT, nrow = 3, padding = unit(0, "line"))
```

Color of the samples were specified with the following codes extracted from `2_gen_radar.R`:
```{r, eval=FALSE}
library(RColorBrewer)
n_member <- nrow(data); n_red_range <- ceiling(n_member/2); n_blue_range <- n_member - n_red_range
# colors in yellow range are avoided for better visualization
colors <- c(colorRampPalette(colors = brewer.pal(11, 'Spectral')[1:4])(n_red_range),
            colorRampPalette(colors = brewer.pal(11, 'Spectral')[8:11])(n_blue_range))
```


## 3. Hierarchical clustering of samples

Using the `gen_dendro()` function, you can cluster the samples using the 5 domains or the sub-categories (argument `cat`); as well as the number of groups (argument `clus_k`). The function generates a matrix of Euclidean distances among samples with `dist()` function then performed hierarchical clustering with `hclust()` using method "complete".
Chunk of codes below is extracted from `3_gen_dendro.R` as demonstration:

```{r, eval=FALSE}
# using sub-categories to cluster samples
cat <- data.frame(member_name, data[,-(1:6)])
dist_cat <- dist(cat[,-1], method = "euclidean")
hc_cat <- hclust(dist_cat, method = "complete")
cluster_cat <- cutree(hc_cat, k = clus_k)
```

Here are the resulting dendrograms generated using `ggdendrogram` package for the domains and their sub-categories. `clus_k` were set to 3 in this example. The shape of the nodes corresponds to the group each sample belongs. You may notice that there are slight differences in the groupings.

```{r echo = FALSE, results = 'axis', fig.align='center'}
library(ggplot2)
load("dendrogram.Rdata")
grid.arrange(dendro_list[[1]] + ggtitle("5 Domains"), dendro_list[[2]] + ggtitle("30 Sub-categories"), nrow = 1)
```

## 4. Bar plots displaying scores of each sub-categories in the domains

With the `gen_bar()` function, bar-plots for each sub-categories of the samples can be visualized for each domains. Bar plots are relatively easy, just a few lines of codes available in `4_gen_bar.R`. Here just to show one of the plots for the domain Neuroticism. Other plots are stored in `bar.Rdata`. Please explore those plots if you are interested.

```{r echo = FALSE, results = 'axis', fig.align='center'}
#library(ggplot2)
load("bar.Rdata")
bar_list[[1]]
```

## 5. Principle Components Analysis (PCA)

PCA plots can be generated with the `gen_PCA()` function. The function uses `prcomp()` to perform PCA using either scores of the domains or the scores of the sub-categories, then generates plots combining the grouping information from the hierarchical clustering by extracting the information from the `prcomp` object. Codes below are extracted from `5_gen_PCA.R` as demonstration:

```{r, eval=FALSE}
load(dendrogram.Rdata)
cat_pca <- prcomp(cat_tbls[[n_cat]][,-1], scale=TRUE)
cluster <- data.frame(member = dendro_list[["cat2"]]$plot_env$ddata_cat$labels$label,
                      cluster = dendro_list[["cat2"]]$plot_env$cluster_cat) %>% arrange(member)
cat_pc_tbl <- data.frame(clusters[[n_cat]], as.data.frame(cat_pca$x)) # construct data.frame combining cluster information and Principle Component (PC) information stored in cat_pca$x
var <- cat_pca$sdev^2 / sum(cat_pca$sdev^2) # extract variance information of all PCs
```

You can set `LoadTF = TRUE` in the `gen_PCA()` function to show the loadings in the plot. The loading points are arranged in descending order of absolute magnitudes in PC1 and PC2 to indicate their relative importance. You can choose the number of top loadings to show in the plot by specifying the `n_load` argument. Code chunk below shows how this part was performed `gen_PCA()`:

```{r, eval=FALSE}
load_tbl <- data.frame(Variables = cat_name, cat_pca$rotation)
top_var <- load_tbl %>% select(Variables, PC1, PC2) %>% 
          mutate(hyp = sqrt(PC1^2+PC2^2)) %>% arrange(desc(hyp)) %>% .[1:n_load,]
```


These are the PCA plots for the 5 domains with and without displaying the loadings.

```{r echo = FALSE, results = 'axis', fig.align='center'}
#library(ggplot2)
load("PCA.Rdata")
PCA_cat1[[1]]
PCA_cat1[[2]]
```

Below are the PCA plots based on the sub-categories without displaying loadings.

```{r echo = FALSE, results = 'axis', fig.align='center'}
#library(ggplot2)
load("PCA.Rdata")
PCA_cat2[[1]]
```

When `LoadTF = TRUE` and `n_load` is set at `10`. The plot looks like this:

```{r echo = FALSE, results = 'axis', fig.align='center'}
#library(ggplot2)
load("PCA.Rdata")
PCA_cat2[[2]][[1]]
```

Each item is shaded according to their big 5 categories. Here is the plot when all items are shown (`n_load` = 30). 

```{r echo = FALSE, results = 'axis', fig.align='center'}
#library(ggplot2)
load("PCA.Rdata")
PCA_cat2[[2]][[5]]
```

You can play around with setting different `n_load` as well.

All the plots and objects created from the sample data are stored in `.Rdata` files. Scripts in `.R` format used to generate the example plots are available in this repo. Enjoy playing around with the data!